---
output:
  pdf_document:
    keep_tex: true
    latex_engine: lualatex
    template: ./eisvogel.tex
    extra_dependencies: ["booktabs"]
    toc: true
title: "January 2020 Parking Violations"
author: "justjensen.co"
date: "`r format(Sys.time(), '%e %b, %Y')`"
subtitle: "Parking Tickets from Washington D.C.'s Open Data Site"
titlepage: true
logo: "justjensen-logo.pdf"
titlepage-rule-color: "6F4A8E"
titlepage-text-color: "000000"
titlepage-rule-height: 2
toc-own-page: true
---

This notebook provides information about parking tickets in Washington D.C.

# Aquiring the Data

First, we'll need to grab data directly in R using the [Parking Ticket dataset]'s API.

```{r, echo=T, results='hide', message=F, warning=F}
# url <- 'https://opendata.arcgis.com/datasets/009dedfbaf364905a8e25181b3490cd9_0.geojson'
# destination_file <- 'january_2020_dc_parking_tickets.geojson'
# download.file(url, destination_file, 'curl')
# library(rgdal)
# sp_parking_tickets <- readOGR('january_2020_dc_parking_tickets.geojson')

library(tidyverse)
url <- 'https://opendata.arcgis.com/datasets/009dedfbaf364905a8e25181b3490cd9_0.csv'
destination_file <- 'january_2020_dc_parking_tickets.csv'
download.file(url, destination_file, 'curl')
df_violations <- read_csv(destination_file)
colnames(df_violations) <- tolower(colnames(df_violations))
```

The first violation in our data set is a '`r df_violations$violation_proc_desc[1]`' violation.

Before going further, we'll need to do a little more work to get our dataframe set up properly!

```{r, echo=T, results='hide', message=F, warning=F}
df_violations <- df_violations[c('objectid', 'issue_date', 'issuing_agency_code',
                                 'issuing_agency_name', 'issuing_agency_short', 'violation_code',
                                 'violation_proc_desc', 'location', 'disposition_type',
                                 'disposition_date', 'fine_amount', 'total_paid', 'latitude', 
                                 'longitude', 'mar_id', 'gis_last_mod_dttm')]
cat(paste('#', 'This is a heading for Object', df_violations$objectid[1], '  \n'))
```

# Creating a Chart of Violations

```{r}
df_violations$issue_date <- as.Date(df_violations$issue_date, format='%Y/%m/%d')
df_violations_per_day <- df_violations %>% count(issue_date)
plt <- ggplot(df_violations_per_day) +
  geom_hline(yintercept=0, size=0.4, color='#3C3C3C')+
  geom_line(aes(x=issue_date, y=n), color='#6f4a8e', alpha=0.8, size=1) +
  geom_point(aes(x=issue_date, y=n), color='#6f4a8e', alpha=0.8, size=2) +
  labs(x='', y='', title='Parking Violations in Washington D.C. drop off on Weekends',
       subtitle="Daily Parking Violations in January 2020 from D.C.'s Open Data Site") +
  scale_x_date(date_labels='%b %d', breaks=seq(as.Date('2020-01-01'), as.Date('2020-01-31'), by='weeks')) +
  theme(text=element_text(size=12, color='#3C3C3C'),
        plot.title=element_text(hjust=0, size=rel(1.5), face='bold'),
        plot.subtitle = element_text(hjust=0, size=rel(1.1)),
        plot.caption=element_text(hjust=0),
        plot.title.position = 'plot',
        plot.background = element_rect(fill='#F0F0F0'), axis.ticks=element_blank(),
        panel.background = element_rect(fill='#F0F0F0'), panel.grid=element_line(color=NULL),
        panel.grid.major=element_line(color='#d2d2d2'), panel.grid.minor=element_blank(),
        strip.background=element_blank(), strip.text=element_text(face='bold'),
        plot.margin=unit(c(1,1,1,1), 'lines'))
print(plt)
ggsave('Parking Violations in Washington DC.png', plot=plt, type='cairo',
       height=5,width=8,units='in')
```

# Creating a Table of Violations

```{r, warning=F, message=F, results='asis'}
library(kableExtra)
df_violations$fine_paid <- ifelse(df_violations$fine_amount==df_violations$total_paid,
                                  1, 0)

df_violations$fine_bin <- case_when(
  df_violations$fine_amount < 50 ~ '<$50',
  df_violations$fine_amount < 100 ~ '$50 - $99',
  df_violations$fine_amount < 200 ~ '$100 - $199',
  df_violations$fine_amount >= 200 ~ '$200+'
)

# Preparing the final dataframe for table generation
df_violations_fines <- df_violations %>%
  drop_na(fine_amount) %>%
  group_by(fine_bin) %>%
  summarise('Tickets (thousands)'=round(length(fine_paid)/1000,1),
            'Percent Paid'=paste0('%',round(sum(fine_paid)/length(fine_paid),3)*100)) %>%
  gather('Tickets (thousands)', 'Percent Paid', key='value_type', value='value') %>%
  spread('fine_bin', 'value') %>%
  select(value_type, '<$50', '$50 - $99', '$100 - $199', '$200+')

colnames(df_violations_fines)[1] <- 'Fine Amount'

kb <- kbl(df_violations_fines, format = 'latex',
          booktabs=T, digits=1, linesep='', align=c('lrrrr')) %>%
  kable_styling(latex_options = 'striped') %>%
  add_header_above(c(' ', 'Cheaper'=2, 'More Expensive'=2))

# kb <- kbl(df_violations_fines,
#             'latex', booktabs=T, digits=2, linesep='',
#             col.names = c('', 'Hours', strftime(recent_weeks[1], '%Y-%m-%d'),
#                           strftime(recent_weeks[2], '%Y-%m-%d'),'Perc. Diff'),
#             align=c('lrrrr')) %>%
#   kable_styling(full_width = F, latex_options = 'striped')
print(kb)
```
